---
title: "Data Visualization"
title-block-banner: TRUE
fig-width: 7
#fig-height: 5
warning: FALSE
error: TRUE
echo: FALSE
message: FALSE
code-fold: TRUE
code-summary: "Reveal Code"
code-copy: hover
---

```{r, Load Data}
#| include: FALSE
library(cpaltemplates)
library(tidycensus)
library(tidyverse)
library(rio)
library(sf)

counties <- tigris::counties(state = "TX")

yearsOld <- lst(2010, 2011)
yearsNew <- lst(2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2021)

acs_varOld <- c(
  tot_pop = "S1703_C01_001", #total population
  pop_u18 = "S1703_C01_004", #population under 18
  pop_bp = "S1703_C03_001", #population below poverty
  bp_u18 = "S1701_C02_002", #population under 18 below poverty
  mhi = "S0501_C01_101", #median household income
  rcb = "S0501_C01_137", #renter cost burden over 30%
  ocb = "S0501_C01_134", #owner cost burden over 30%
  rohh = "S0501_C01_135", #renter occupied households
  oohh = "S0501_C01_132" #owner occupied households
)

acs_varNew <- c(
  tot_pop = "S1701_C01_001", #total population
  pop_u18 = "S1701_C01_002", #population under 18
  pop_bp = "S1701_C02_001", #population below poverty
  bp_u18 = "S1701_C02_002", #population under 18 below poverty
  mhi = "S0501_C01_101", #median household income
  rcb = "S0501_C01_137", #renter cost burden over 30%
  ocb = "S0501_C01_134", #owner cost burden over 30%
  rohh = "S0501_C01_135", #renter occupied households
  oohh = "S0501_C01_132" #owner occupied households
)

##### Pull Data from 2010 to 2020 #####
cpal_old <- map(
  yearsOld,
  ~get_acs(
    geography = "place", 
    variables = acs_varOld,
    year = .x,
    survey = "acs1", 
    output = "wide"),
  ) %>%
  map2(yearsOld, ~mutate(.x, year = .y)) %>%
  reduce(., rbind) %>%
  mutate(NAME = str_replace(NAME, " city", "")) %>%
  mutate(cbp = bp_u18E/pop_u18E,
         bp = pop_bpE/100,
         pop_bpE = bp*tot_popE,
         #bp_u18E = cbp*pop_u18E,
         pop_bpM = round(pop_bpE*(pop_bpM/100), digits = 0),
         #bp_u18M = round(bp_u18E*(bp_u18M/100), digits = 0),
         pop_bpE = round(pop_bpE, digits = 0),
         #bp_u18E = round(bp_u18E, digits = 0)
  )

cpal_2020 <- get_acs(
    geography = "place", 
    variables = acs_varNew,
    year = 2020, 
    survey = "acs5", 
    output = "wide") %>%
  mutate(cbp = round(bp_u18E/pop_u18E, digits = 4),
         bp = round(pop_bpE/tot_popE, digits = 4),
         NAME = str_replace(NAME, " city", ""),
         year = 2020)

cpal_multi <- map(
  yearsNew,
  ~get_acs(
    geography = "place", 
    variables = acs_varNew,
    year = .x, 
    survey = "acs1", 
    output = "wide"),
  ) %>%
  map2(yearsNew, ~mutate(.x, year = .y)) %>%
  reduce(., rbind) %>%
  mutate(cbp = round(bp_u18E/pop_u18E, digits = 4),
         bp = round(pop_bpE/tot_popE, digits = 4),
         NAME = str_replace(NAME, " city", ""))

cpal_dallas <- get_acs(
    geography = "tract", 
    state = "TX",
    county = "Dallas County",
    variables = acs_varNew,
    year = 2021, 
    survey = "acs5", 
    output = "wide") %>%
  mutate(cbp = round(bp_u18E/pop_u18E, digits = 4),
         bp = round(pop_bpE/tot_popE, digits = 4))

##### Join into Complete Dataset #####
cpal_comp <- rbind(cpal_old, cpal_multi) %>%
  rbind(., cpal_2020) %>%
  group_by(NAME) %>%
  arrange(NAME, year) %>%
  mutate(rcbE = rcbE/100,
         ocbE = ocbE/100) %>%
  mutate(cbp_diff = cbp-lag(cbp, n=1),
         cbp_ave = (cbp+lag(cbp, n=1)+lag(cbp, n=2))/3,
         cbp_avedif = cbp_ave-lag(cbp_ave, n=1),
         
         rcb_diff = rcbE-lag(rcbE, n=1),
         rcb_ave = (rcbE+lag(rcbE, n=1)+lag(rcbE, n=2))/3,
         rcb_avedif = rcb_ave-lag(rcb_ave, n=1),
         
         ocb_diff = ocbE-lag(ocbE, n=1),
         ocb_ave = (ocbE+lag(ocbE, n=1)+lag(ocbE, n=2))/3,
         ocb_avedif = ocb_ave-lag(ocb_ave, n=1)) %>%
  ungroup() %>%
  mutate(highlight=ifelse(NAME == "Dallas, Texas", "DAL", "NONE"),
         year = as.Date(paste0(year, "-01-01")))

##### Filter out multiple groups based on population size #####
cpal_1mil <- cpal_comp %>%
  group_by(NAME) %>%
  filter(any(tot_popE >= 1000000)) %>%
  mutate(CODE = ifelse(NAME == "Chicago, Illinois", "CHI", 
                       ifelse(NAME == "Dallas, Texas", "DAL", 
                              ifelse(NAME == "Houston, Texas", "HOU", 
                                     ifelse(NAME == "Los Angeles, California", "LA", 
                                            ifelse(NAME == "New York, New York", "NYC", 
                                                   ifelse(NAME == "Philadelphia, Pennsylvania", "PHI", 
                                                          ifelse(NAME == "Phoenix, Arizona", "PHO", 
                                                                 ifelse(NAME == "San Antonio, Texas", "SATX", 
                                                                        ifelse(NAME == "San Diego, California", "SD", 
                                                                               ifelse(NAME == "San Jose, California", "SJ", "ERROR")))))))))))
cpal_texas <- cpal_comp %>%
  filter(str_detect(NAME, ", Texas")) %>%
  group_by(NAME) %>%
  filter(any(tot_popE >= 500000)) %>%
  mutate(CITY = str_remove(NAME, ", Texas"))

rm(cpal_old)
rm(cpal_multi)
rm(cpal_2020)
rm(yearsNew)
rm(yearsOld)

```

# Background
If you have any questions, please contact Michael Lopez (michael@childpovertyactionlab.org).

## Graphic Design Principles

## Installing Templates

## Tips and Tricks

# Scatter Plots

## One-Color Scatter PLot
Scatter plots are useful for visualizing the relationship between two variables, identifying patterns or trends, and detecting outliers or anomalies.
```{r, One Color Scatter Plot}
#| echo: TRUE
#| fig-cap: A basic scatterplot visualizing a relationship between Child Poverty Rate and Median Household Income in cities with over a million population.
ggplot(cpal_1mil, aes(x=cbp, y=mhiE)) + 
  geom_point(size=2, color = "#008097") +
  theme_cpal() +
  labs(title = "MHI vs Child Poverty Rate",
       subtitle = "Cities Over 1,000,000 Population",
       x = "Child Poverty Rate",
       y = "Median Household Income") +
  scale_y_continuous(labels = scales::dollar_format(), limits = c(0, NA)) + # <1> Format scale to $ amounts and start at 0.
  scale_x_continuous(labels = scales::percent_format()) # <2> Format scale into percentage.
```

```{r, High Density Scatter Plot}
#| echo: TRUE
#| fig-cap: A high density scatterplot visualizing a relationship between Child Poverty Rate and Median Household Income in cities with over a million population.
ggplot(cpal_comp, aes(x=cbp, y=mhiE)) + 
  geom_point(size=1, alpha = 0.2, color = "#008097") + #<1> alpha changes the transparency of points plotted.
  theme_cpal() +
  labs(title = "MHI vs Child Poverty Rate",
       subtitle = "Cities in the United States",
       x = "Child Poverty Rate",
       y = "Median Household Income") +
  scale_y_continuous(labels = scales::dollar_format(), limits = c(0, NA)) +
  scale_x_continuous(labels = scales::percent_format())
```

## Verying Point Size Scatter Plot
```{r, Varying Point Size Scatter Plot}
#| echo: TRUE
#| fig-cap: A size adjusted scatterplot visualizing a relationship between Child Poverty Rate and Median Household Income in cities with over a million population with points adjusted based on total number of children below poverty.

BAMMtools::getJenksBreaks(cpal_1mil$bp_u18E, k = 4)

cpal_1mil%>%
  ggplot(aes(x=cbp, y=mhiE)) + 
  geom_point(aes(size = pop_u18E), alpha = 0.3, color = "#008097") + # <1> size aesthetic can be mapped to a variable in dataframe, but scale_radius will need to be defined
  scale_radius(range = c(3, 15), 
               breaks = c(200000, 430000, 870000, 1800000),
               labels = scales::comma) +
  theme_cpal() +
  labs(title = "MHI vs Child Poverty Rate",
       subtitle = "Cities in the United States",
       x = "Child Poverty Rate",
       y = "Median Household Income") +
  scale_y_continuous(labels = scales::dollar_format(), limits = c(0, NA)) +
  scale_x_continuous(labels = scales::percent_format())
```

# Line Graph

## Basic Line Graph
```{r, Line Graph}
#| echo: TRUE
#| fig-cap: A basic line graph visualizing a the Child Poverty Rate in the City of Dallas
unique(cpal_1mil$CODE)

cpal_1mil %>%
  filter(highlight == "DAL") %>%
  ggplot(aes(x=year, y=cbp, group = CODE, color = CODE)) + 
  geom_line(size = 2) +
  geom_point(size=3) +
  theme_cpal() +
  labs(title = "Child Poverty Rate in the City of Dallas",
       subtitle = "Between 2010 to 2021",
       x = "Year",
       y = "Child Poverty Rate") +
  scale_color_cpal(theme = "factor") +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, NA)) +
  scale_x_date(date_breaks = "3 years", date_labels = "%Y")
```

## Multiples Line Graph
```{r, Multiple Line Graph}
#| echo: TRUE
#| fig-cap: A multiples line graph visualizing the child poverty rate of multiple cities since 2010.
cpal_1mil %>%
    filter(CODE %in% c("DAL", "HOU", "SATX")) %>%
  ggplot(aes(x=year, y=cbp, group = CODE, color = CODE)) + 
  geom_line(size = 2) +
  geom_point(size=3) +
  theme_cpal() +
  labs(title = "Child Poverty Rate in Texas Cities",
       subtitle = "Between 2010 to 2021",
       x = "Year",
       y = "Child Poverty Rate") +
  scale_color_cpal(theme = "factor") +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, NA)) +
  scale_x_date(date_breaks = "1 years", date_labels = "%Y")
```

## Bar Graph
```{r, Bar Graph}
#| echo: TRUE
#| fig-cap: A basic line graph visualizing a the Child Poverty Rate in the City of Dallas
# A basic line graph visualizing a the Child Poverty Rate in the City of Dallas
cpal_1mil %>%
  filter(year == as.Date("2021-01-01")) %>%
  ggplot(aes(x=cbp, y=CODE)) + 
  geom_bar(stat = "identity", size = 2, fill = "#008097", color = "#E7ECEE") +
  theme_cpal() +
  labs(title = "Child Poverty Rate in the City of Dallas",
       x = "Year",
       y = "Child Poverty Rate") +
  scale_color_cpal(theme = "factor") +
  scale_x_continuous(labels = scales::percent_format(), limits = c(0, NA))
```

## Histogram
```{r, Histogram}
#| echo: TRUE
#| fig-cap: A basic histogram visualizing 
# A basic line graph visualizing a the Child Poverty Rate in the City of Dallas
cpal_dallas %>%
  ggplot(aes(x = pop_u18E)) + 
  geom_histogram(colour="#E7ECEE", fill="#008097") +
  geom_vline(aes(xintercept = mean(pop_u18E)), color="#E98816", linetype="dashed", size=1) +
  theme_cpal() +
  labs(title = "Histogram of Children Below Poverty in City of Dallas",
       x = "Year",
       y = "Child Poverty Rate")
```
## Facet Graph

## Map

# Interactive Graphics

## plotly

## leaflet
```{r, Leaflet}
#| echo: TRUE
# A basic leaflet map

#counties <- tigris::counties(state = "TX")

dallas <- filter(counties, NAME == "Dallas")

bbox <- dallas %>%
      st_bbox(.) %>%
      as.vector()

leaflet(dallas) %>%
  fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>%
  addTiles(urlTemplate = cpal_mapbox, attribution = cpal_leaflet) %>%
  addPolygons(weight = 1,     
              opacity = 0.5,
              color = "#686158",
              fillOpacity = 0.1)

```
