---
title: "Packages and Functions"
title-block-banner: TRUE
fig-width: 7
fig-height: 5
warning: FALSE
error: TRUE
echo: FALSE
message: FALSE
---

## cpaldirectory
Directory for CPAL Analytics team to organize visualization code and projects into a centralized location. This website is built from this repository.

You can install the latest version of `cpaldirectory` from GitHub:

```{r}
# | echo = TRUE

# install.packages("devtools")
#devtools::install_github("childpovertyactionlab/cpaldirectory")
```

## cpaltemplates

Primary package containing templates and tools for new project generation across CPAL Analytics data projects. Contains templates for shiny, web reports, websites, project directories, and visualizations.

You can install the latest version of `cpaltemplates` from GitHub:

```{r}
# | echo = TRUE

# install.packages("devtools")
#devtools::install_github("childpovertyactionlab/cpaltemplates")
```

## CPALtools 
Soon to be depreciated and replaced with `cpaltools`. A package containing functions for analysis by CPAL Analytics team.

You can install the latest version of `CPALtools` from GitHub:

```{r}
# | echo = TRUE

# install.packages("devtools")
#devtools::install_github("childpovertyactionlab/CPALtools")
```
